TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-09 paper/draft.tex
add software refs
scipy ref
reference table 5
start expanding discussion

\newcommand{\apogee}{\textsl{APOGEE}}
\newcommand{\thecannon}{\textsl{The Cannon}}
\newcommand{\cannon}{\textsl{Cannon}}
\newcommand{\aspcap}{\textsl{ASPCAP}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\zmass}{\textsl{2MASS}}
\newcommand{\sdss}{\textsl{SDSS}}
\newcommand{\kepler}{\textsl{Kepler}}
\newcommand{\lamost}{\textsl{LAMOST}}
\newcommand{\sdssiv}{\textsl{SDSS-IV}}
\newcommand{\tess}{\textsl{TESS}}
\newcommand{\msun}{M_{\odot}}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\logg}{\log g}
\newcommand{\feh}{[{\mathrm{Fe}/\mathrm{H}}]}

@@

\subsection{Masses \& Radii}
Masses and radii for our test sample are estimated using photometric empirical relations from \citealt{Mann:2019} and \citealt{Mann:2015} respectively.

@@

% % The scatter of each relation being largely due to metallicity variation, both for the \cannon\ and \aspcap\ metallicities.
% \cannon\ temperatures show the highest agreement with the photometric relations from \citealt{Mann:2015} and \citealt{Boyajian:2012}, with a scatter of 80K.

@@

\begin{enumerate}
\item[-] Brief summary of what we did and found

\item[-] Things we learned along the way: For example, not all training sets worked well in our experiments, probably because they were too small or had too little dynamic range.

\item[-] \color{gcolor}{HOGG: Some discussion of precision and accuracy: \thecannon\ can produce precise results, but is only as accurate as its training set.}\color{black}

\item[-] Some discussion of whether the derivatives are sensible or surprising. \color{gcolor}{HOGG: I'm not sure what to say here}\color{black}

\item[-] \color{gcolor}{HOGG:\thecannon\ as a test of evaluating training parameter accuracy?}\color{black}

\item[-] \color{gcolor}{HOGG: Discussion of validation tests and ground truth}\color{black}

\end{enumerate}

Future work:
\begin{enumerate}
\item[-] Create a binary sample; using color/magnitude \citealt{Widmark:2018} and by spectroscopy \citealt{ElBadry:2018}

\item[-] \color{gcolor}{HOGG: Discussion of the implications of these results for the NASA \tess\ Mission and how the future might change..?}\color{black}
\end{enumerate}

@@

@@

\begin{deluxetable*}{rcl}
\tablecaption{Test results table header description.}
\tablehead{Column & Unit & Description}
\startdata
{\tt APOGEE$\_$ID} 	& 	& \apogee\ \zmass\ designation \\
{\tt GAIA$\_$ID} 		&	& \gaia\ identification number \\
{\tt TEFF} 				& K	& Mann-trained \cannon\ effective temperature	\\
{\tt FE$\_$H }		& dex	& Mann-trained \cannon\ $\feh$	\\
{\tt SPT} 				& 	& West-trained \cannon\ spectral subtype (M0$-$M9)	\\
{\tt CHI$\_$MANN} 	& 	& $\chi^2$ fit of Mann-trained \cannon\ model	\\
{\tt CHI$\_$WEST} 	& 	& $\chi^2$ fit of West-trained \cannon\ model	\\
{\tt RADIUS} 			& R$_\odot$	& Radius determined by \citealt{Mann:2015} relation	\\
{\tt MASS} 			& M$_\odot$	& Mass determined by \citealt{Mann:2019} relation	\\
{\tt J} 				& mag	& \apogee\ $J$ band photometry	\\
{\tt J$\_$ERR} 		& mag	& \apogee\ $J$ band photometry uncertainty	\\
{\tt H} 				& mag	& \apogee\ $H$ band photometry	\\
{\tt H$\_$ERR} 		& mag	& \apogee\ $H$ band photometry uncertainty	\\
{\tt K} 				& mag	& \apogee\ $K$ band photometry	\\
{\tt K$\_$ERR} 		& mag	& \apogee\ $K$ band photometry uncertainty	\\
{\tt BP$\_$MAG }		& mag	& \gaia\ $BP$ band photometry	\\
{\tt RP$\_$MAG} 		& mag	& \gaia\ $RP$ band photometry	\\
{\tt G$\_$MAG} 		& mag	& \gaia\ $G$ band photometry	\\
{\tt G$\_$ABS} 		& mag	& \gaia\ $G$ band absolute magnitude	\\
{\tt BP$\_$RP} 		& mag	& \gaia\ $BP-RP$ color	\\
{\tt RA} 				& deg	& \apogee\ radial ascension angle	\\
{\tt DEC} 			& deg	& \apogee\ declination angle	\\
{\tt PMRA} 			& mas yr$^{-1}$	& \gaia\ radial ascension proper motion	\\
{\tt PMRA$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ radial ascension proper motion uncertainty	\\
{\tt PMDEC} 			& mas yr$^{-1}$	& \gaia\ declination proper motion	\\
{\tt PMDEC$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ declination proper motion uncertainty	\\
{\tt PLX} 			& mas	& \gaia\ parallax	\\
{\tt PLX$\_$ERR} 		& mas	& \gaia\ parallax uncertainty	\\
{\tt DIST} 			& pc	& Distance (1/$\varpi$)	\\
{\tt RV$\_$APOGEE} 	& km s$^{-1}$	& \apogee\ radial velocity	\\
{\tt RV$\_$APOGEE$\_$ERR} 	& km s$^{-1}$	& \apogee\ radial velocity uncertainty	\\
{\tt RV$\_$GAIA} 				& km s$^{-1}$	& \gaia\ radial velocity	\\
{\tt RV$\_$GAIA$\_$ERR}		& km s$^{-1}$	& \gaia\ radial velocity uncertainty	\\
{\tt Vx} 				& km s$^{-1}$	& Cartesian $x$ velocity in Galactocentric coordnates	\\
{\tt Vy} 				& km s$^{-1}$	& Cartesian $y$ velocity in Galactocentric coordnates	\\
{\tt Vz} 				& km s$^{-1}$	& Cartesian $z$ velocity in Galactocentric coordnates	\\
{\tt X} 				& pc	& Cartesian $x$ position in Galactocentric coordnates	\\
{\tt Y} 				& pc	& Cartesian $y$ position in Galactocentric coordnates	\\
{\tt Z} 				& pc	& Cartesian $z$ position in Galactocentric coordnates	
\enddata
\end{deluxetable*}

>>>>

\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\apogee}{\textsl{APOGEE}}
\newcommand{\thecannon}{\textsl{The Cannon}}
\newcommand{\cannon}{\textsl{Cannon}}
\newcommand{\aspcap}{\textsl{ASPCAP}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\zmass}{\textsl{2MASS}}
\newcommand{\sdss}{\textsl{SDSS}}
\newcommand{\kepler}{\textsl{Kepler}}
\newcommand{\lamost}{\textsl{LAMOST}}
\newcommand{\sdssiv}{\textsl{SDSS-IV}}
\newcommand{\tess}{\textsl{TESS}}
\newcommand{\msun}{M_{\odot}}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\logg}{\log g}
\newcommand{\feh}{[{\mathrm{Fe}/\mathrm{H}}]}

@@

Table \ref{table:test_results} presents...

@@

@@

We trained and evaluated the performance of \thecannon\ at deriving atmospheric temperatures and metallicities from high resolution near-infrared spectra... summarize results.

While \thecannon\ generally achieves excellent precision at predicting labels and reproducing the spectral features as demonstrated in this study and previous work on \thecannon\ (\citealt{Ness:2015}, \citealt{Casey:2016}), the accuracy of labels from model application to new sources is limited by the accuracy, relative precision, size, dynamic range and representation of the training sample.

Comment on the importance of each property of training sample: \\
accuracy -  \\
precision -  \\
size -  \\
dynamic range - training sample must span \\
representation - \\

Talk about the particular limitations we have found in this training sample.

% \begin{enumerate}
% \item[-] Brief summary of what we did and found

% \item[-] Things we learned along the way: For example, not all training sets worked well in our experiments, probably because they were too small or had too little dynamic range.

% \item[-] \color{gcolor}{HOGG: Some discussion of precision and accuracy: \thecannon\ can produce precise results, but is only as accurate as its training set.}\color{black}

% \item[-] Some discussion of whether the derivatives are sensible or surprising.

% \item[-] \color{gcolor}{HOGG:\thecannon\ as a test of evaluating training parameter accuracy?}\color{black}

% \item[-] \color{gcolor}{HOGG: Discussion of validation tests and ground truth}\color{black}
% \end{enumerate}

% Future work:
% \begin{enumerate}
% \item[-] Create a binary sample; using color/magnitude \citealt{Widmark:2018} and by spectroscopy \citealt{ElBadry:2018}

% \item[-] \color{gcolor}{HOGG: Discussion of the implications of these results for the NASA \tess\ Mission and how the future might change..?}\color{black}
% \end{enumerate}

@@

\facilities{
	\sdssiv(\apogee),
	\gaia}

\software{
	\code{Astropy} \citep{astropy,astropy2}, \,
	\code{matplotlib} \citep{Hunter:2007}, \,
	\code{numpy} \citep{vanderWalt:2011}, \,
	\code{scipy} \citep{Jones:2001}, \,
	\code{The Cannon} \citep{Ness:2015}}

@@

\begin{deluxetable*}{rcl}
\tablecaption{Test results table header description.}
\tablehead{Column & Unit & Description}
	\startdata
	{\tt APOGEE$\_$ID} 	& 	& \apogee\ \zmass\ designation \\
	{\tt GAIA$\_$ID} 		&	& \gaia\ identification number \\
	{\tt TEFF} 				& K	& Mann-trained \cannon\ effective temperature	\\
	{\tt FE$\_$H }		& dex	& Mann-trained \cannon\ $\feh$	\\
	{\tt SPT} 				& 	& West-trained \cannon\ spectral subtype (M0$-$M9)	\\
	{\tt CHI$\_$MANN} 	& 	& $\chi^2$ fit of Mann-trained \cannon\ model	\\
	{\tt CHI$\_$WEST} 	& 	& $\chi^2$ fit of West-trained \cannon\ model	\\
	{\tt RADIUS} 			& R$_\odot$	& Radius determined by \citealt{Mann:2015} relation	\\
	{\tt MASS} 			& M$_\odot$	& Mass determined by \citealt{Mann:2019} relation	\\
	{\tt J} 				& mag	& \apogee\ $J$ band photometry	\\
	{\tt J$\_$ERR} 		& mag	& \apogee\ $J$ band photometry uncertainty	\\
	{\tt H} 				& mag	& \apogee\ $H$ band photometry	\\
	{\tt H$\_$ERR} 		& mag	& \apogee\ $H$ band photometry uncertainty	\\
	{\tt K} 				& mag	& \apogee\ $K$ band photometry	\\
	{\tt K$\_$ERR} 		& mag	& \apogee\ $K$ band photometry uncertainty	\\
	{\tt BP$\_$MAG }		& mag	& \gaia\ $BP$ band photometry	\\
	{\tt RP$\_$MAG} 		& mag	& \gaia\ $RP$ band photometry	\\
	{\tt G$\_$MAG} 		& mag	& \gaia\ $G$ band photometry	\\
	{\tt G$\_$ABS} 		& mag	& \gaia\ $G$ band absolute magnitude	\\
	{\tt BP$\_$RP} 		& mag	& \gaia\ $BP-RP$ color	\\
	{\tt RA} 				& deg	& \apogee\ radial ascension angle	\\
	{\tt DEC} 			& deg	& \apogee\ declination angle	\\
	{\tt PMRA} 			& mas yr$^{-1}$	& \gaia\ radial ascension proper motion	\\
	{\tt PMRA$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ radial ascension proper motion uncertainty	\\
	{\tt PMDEC} 			& mas yr$^{-1}$	& \gaia\ declination proper motion	\\
	{\tt PMDEC$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ declination proper motion uncertainty	\\
	{\tt PLX} 			& mas	& \gaia\ parallax	\\
	{\tt PLX$\_$ERR} 		& mas	& \gaia\ parallax uncertainty	\\
	{\tt DIST} 			& pc	& Distance (1/$\varpi$)	\\
	{\tt RV$\_$APOGEE} 	& km s$^{-1}$	& \apogee\ radial velocity	\\
	{\tt RV$\_$APOGEE$\_$ERR} 	& km s$^{-1}$	& \apogee\ radial velocity uncertainty	\\
	{\tt RV$\_$GAIA} 				& km s$^{-1}$	& \gaia\ radial velocity	\\
	{\tt RV$\_$GAIA$\_$ERR}		& km s$^{-1}$	& \gaia\ radial velocity uncertainty	\\
	{\tt Vx} 				& km s$^{-1}$	& Cartesian $x$ velocity in Galactocentric coordnates	\\
	{\tt Vy} 				& km s$^{-1}$	& Cartesian $y$ velocity in Galactocentric coordnates	\\
	{\tt Vz} 				& km s$^{-1}$	& Cartesian $z$ velocity in Galactocentric coordnates	\\
	{\tt X} 				& pc	& Cartesian $x$ position in Galactocentric coordnates	\\
	{\tt Y} 				& pc	& Cartesian $y$ position in Galactocentric coordnates	\\
	{\tt Z} 				& pc	& Cartesian $z$ position in Galactocentric coordnates	
\enddata
\end{deluxetable*} \label{table:test_results}
